---
title: "Baseline Study for the Tuungane Health and Conservation Project"
subtitle: "Social Assessment of Lake Tanganyika Households"
author: 
  - "Sustain East Africa"
date: "March 2024"
output: 
  sea::sea_paged:
    draft_watermark: true
    remove_footer_title: true
    number_sections: false
    front_cover:
      - "img/image_cover3.jpg"
bibliography: references.bib  
biblio-style: "apalike"
nocite: '@*'
toc-title: "Table of Contents"
lof: true
lot: true
lof-unlisted: true
lot-unlisted: true
knit: pagedown::chrome_print
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
library(sea)
library(tidyverse)
library(xfun)
library(lubridate)
library(openxlsx)
library(scales)
library(googlesheets4)
library(stats)
library(stats4)
library(survey)
library(srvyr, warn.conflicts = FALSE)
library(cowplot)
library(egg)
library(tibble)
library(fontawesome)
library(sea)

#Sustain EA colour palette
SEA_palette <- c("#d77e5e", "#a4b792", "#e6e7e2", "#3d5919", "#202C39", "#381D2A", "#000000")

# Load data
tanganyika_clean <- readRDS("tanganyika_clean.rds")
demo <- readRDS("survey_demo.rds")

# Set survey design
tanganyika_survey <- tanganyika_clean %>%
    as_survey_design(strata = stype, fpc = fpc)
```

### Disclaimer

Disclaimer

### Rights

Rights

### Suggested citation

Please cite this report as such :

### Acknowledgements

Acknowledgements

<!-- inverse gunmetal page -->

::: {.section-gunmetal-inverse}
# Summary

## Context

Provide breif overview of context.

## Key findings

### Positive social impacts

Provide Broad overview of key findings

:::

# Introduction

Introduction

</div>

<!-- rust page -->
<div class ="section-rust">

<!-- quote -->
> This is how to do a quote


```{r}
sea::add_lot_link(caption = 'Caption of table with flextable')

mtcars %>%
  head(7) %>%
  sea::sea_table(
    dark_color = sea_colors("rust_orange"),
    pale_color = sea_colors("grey"),
    third_color = "white"
  ) |>
  flextable::autofit()
```
:::
:::

</article>

<!-- inverse rust page -->

::: {.section-rust-inverse}
# Findings

## Characteristics of the respondents

### Household head's gender, age, and average number of children

Highlight findings

:::

<!-- rust page -->

<div class="section-rust">

# Population Pyramid
```{r, fig.cap='Population pyramid', fig.align='center', out.width = '70%'}
age_levels <- c("<5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34",
                "35-39", "40-44", "45-49", "50-54", "55-59", "60-64",
                "65-69", "70-74", "75-79", "80+")

# Categorize age groups and ensure they have the correct order
demo <- demo %>%
  mutate(age_group = case_when(
    age >= 80 ~ "80+",
    age >= 75 & age < 80 ~ "75-79",
    age >= 70 & age < 75 ~ "70-74",
    age >= 65 & age < 70 ~ "65-69",
    age >= 60 & age < 65 ~ "60-64",
    age >= 55 & age < 60 ~ "55-59",
    age >= 50 & age < 55 ~ "50-54",
    age >= 45 & age < 50 ~ "45-49",
    age >= 40 & age < 45 ~ "40-44",
    age >= 35 & age < 40 ~ "35-39",
    age >= 30 & age < 35 ~ "30-34",
    age >= 25 & age < 30 ~ "25-29",
    age >= 20 & age < 25 ~ "20-24",
    age >= 15 & age < 20 ~ "15-19",
    age >= 10 & age < 15 ~ "10-14",
    age >= 5 & age < 10 ~ "5-9",
    age < 5 ~ "<5"
  )) %>%
  mutate(age_group = factor(age_group, levels = age_levels))  # Set factor levels for ordering

# Calculate overall percentages by age group and gender, excluding unwanted sex value
total_count <- nrow(demo %>% filter(sex != "I DO NOT WANT TO ANSWER"))  # Total population count for reference

demo_summary <- demo %>%
  filter(sex != "I DO NOT WANT TO ANSWER") %>%
  count(age_group, sex) %>%
  mutate(percentage = n / total_count * 100)  # Calculate percentage of entire population

# Plot population pyramid
ggplot(demo_summary, aes(x = age_group, y = ifelse(sex == "MALE", -percentage, percentage), fill = sex)) +
  geom_bar(stat = "identity", width = 0.8) +
  coord_flip() +
  scale_y_continuous(labels = abs, limits = c(-10, 10)) +
  labs(title = "Population Pyramid", x = "Age Group", y = "Percentage of Total Population") +
  scale_fill_manual(values = c("MALE" =  "#3d5919", "FEMALE" =  "#a4b792")) +
  theme_minimal()
```

:::

```{r,  fig.cap='Main type of water treatment in the dry season', fig.align='center', out.width = '70%'}

treatment_dry_expanded <- tanganyika_clean %>% select(hh_code, village, treatment_method_dry, stype, fpc) %>%
  separate_rows(treatment_method_dry, sep = "\\|") %>%
  mutate(treatment_method_dry = trimws(treatment_method_dry)) %>% drop_na()  

treatment_dry_design <- treatment_dry_expanded %>%
  as_survey_design(strata = stype, fpc = fpc)

treatment_data <- treatment_dry_design %>%
  group_by(village, treatment_method_dry) %>%
  summarise(proportion = survey_mean(vartype = "ci", na.rm = TRUE), total = survey_total(vartype = "ci", na.rm = TRUE), n = unweighted(n())) %>% ungroup()

ggplot(treatment_data, aes(x = village, y = proportion * 100, group = treatment_method_dry, fill = treatment_method_dry)) +
  geom_bar(stat = "identity", position = position_dodge(preserve = "single"), width = 0.95) +
  geom_errorbar(aes(ymax = pmin(proportion_upp * 100, 100), ymin = pmax(proportion_low * 100, 0)),
                position = position_dodge(preserve = "single", width = 0.95), width = 0.1) +
  guides(fill = guide_legend(title = NULL)) +
  scale_fill_manual(values = c(SEA_palette)) +
  labs(x = "Village", y = "Percentage of Households") +
  sea::scale_fill_sea_discrete() +
  sea::theme_sea() + 
  theme(legend.position = "top", legend.direction = "horizontal", legend.justification = "center", axis.text.x = element_text(angle = 45, hjust = 1))


```


Self-assessment of the ability to meet daily needs at village level

```{r,  fig.cap='Self-assessment of the ability to meet daily needs at village level', fig.align='center', out.width = '70%'}

village_data <- tanganyika_survey %>%
  group_by(village, household_ability) %>%
  summarise(proportion = survey_mean(vartype = "ci", na.rm = TRUE), total = survey_total(vartype = "ci", na.rm = TRUE), n = unweighted(n()))
  
ggplot(village_data, aes(x = village, y = proportion*100, group = household_ability, fill = household_ability)) +
  geom_bar(stat = "identity", position = position_dodge(preserve = "single"), width = 0.95) +
  geom_errorbar(
    aes(ymax = pmin(proportion_upp, 1), ymin = pmax(proportion_low, 0)),
    position = position_dodge(preserve = "single", width = 0.95), width = 0.1) +
  guides(fill = guide_legend(title = NULL)) +
  scale_fill_manual(values = c(SEA_palette)) +
  labs(x = "Village", y = "Percentage of Households") +
  sea::scale_fill_sea_discrete() +
  sea::theme_sea() + 
  theme(legend.position = "top", legend.direction = "horizontal", legend.justification = "center", axis.text.x = element_text(angle = 45, hjust = 1))


```

Type of fishing boats used at village level

```{r}

```

Household items

```{r,  fig.cap='Household item ownership', fig.align='center', out.width = '70%'}
household_item_expanded <- tanganyika_clean %>% select(hh_code, village, household_item, stype, fpc) %>%
  separate_rows(household_item, sep = "\\|") %>%
  mutate(household_item = trimws(household_item)) %>% drop_na()  

household_item_design <- household_item_expanded %>%
  as_survey_design(ids = hh_code, strata = stype, fpc = fpc, nest = TRUE)

household_item_data <- household_item_design %>%
  group_by(household_item) %>%
  summarise(
    proportion = survey_mean(vartype = "ci", na.rm = TRUE),
    total = survey_total(vartype = "ci", na.rm = TRUE),
    n = unweighted(n())) %>% ungroup()

# Create the plot
ggplot(household_item_data, aes(x = household_item, y = proportion * 100, fill = household_item)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_errorbar(aes(ymax = pmin(proportion_upp, 1), ymin = pmax(proportion_low, 0)),
                width = 0.2) +
  guides(fill = guide_legend(title = NULL)) +
  labs(x = "Household Items", y = "Percentage of Household Ownership") +
  sea::scale_fill_sea_discrete() +
  sea::theme_sea() + 
  theme(legend.position = "top", legend.direction = "horizontal", legend.justification = "center", axis.text.x = element_text(angle = 45, hjust = 1))


```

Proportion of households that borrowed money in the last year at village level
```{r}

```

Purpose of the loan

```{r}

```

Source of loans

```{r}

```

Reason for not having borrowed any money in the previous year

```{r}

```

Relative importance of different species at village level

```{r}

```


### Principal livelihoods of the households

Residents surrounding
### Principal livelihoods of the households

Residents surrounding


```{r}

# Define plotting and summary table function for a single survey variable
hh_info_survey_variable <- function(variable_name, data = tanganyika_clean) {
  # Set survey design
  strat_design <- data %>%
    as_survey_design(strata = stype, fpc = fpc, variables = c(stype, fpc, village, !!sym(variable_name)))
  
  # Group by village and the specified variable, then calculate proportions and totals
  village_data <- strat_design %>%
    group_by(village, !!sym(variable_name)) %>%
    summarise(
      proportion = survey_mean(vartype = "ci", na.rm = TRUE), 
      total = survey_total(vartype = "ci", na.rm = TRUE), 
      n = unweighted(n())
    )
  
  # Prepare summary table
  summary_table <- village_data %>%
    select(village, !!sym(variable_name), n, total, total_low, total_upp, proportion, proportion_low, proportion_upp) %>%
    rename(
      Total = total, 
      Lower_Total_CI = total_low, 
      Upper_Total_CI = total_upp, 
      Proportion = proportion, 
      Lower_Proportion_CI = proportion_low, 
      Upper_Proportion_CI = proportion_upp
    )
  
  # Save summary table as CSV
  table_file_name <- here::here("images", paste0(variable_name, "_summary_table.csv"))
  write.csv(summary_table, table_file_name, row.names = FALSE)
  
  # Create plot
  plot <- ggplot(village_data, aes(x = village, y = proportion, group = !!sym(variable_name), fill = !!sym(variable_name))) +
    geom_bar(stat = "identity", position = position_dodge(preserve = "single"), width = 0.95) +
    geom_errorbar(aes(ymax = pmin(proportion_upp, 1), ymin = pmax(proportion_low, 0)),
                  position = position_dodge(preserve = "single", width = 0.95), width = 0.1) +
    guides(fill = guide_legend(title = NULL)) +
    scale_fill_manual(values = c("#A9CCE3", "#2E86C1", "#F5B7B1", "#D091BB", "#BBD4A6", "#FAD7A0", "#DFDFDF", SEA_palette)) +
    labs(
      title = paste("Proportion of", variable_name, "by Village"),
      x = "Village", 
      y = "Proportion of Households"
    ) +
    scale_y_continuous(limits = c(0, 1)) +
    theme_minimal()
  
  # Return both plot and summary table
  list(plot = plot, table = summary_table)
}

# Plot and summary table for born_ward variable
output <- hh_info_survey_variable("born_ward")
output$plot  # Display plot
output$table  # Display summary table if needed
```


<!-- putting bibliography before appendix -->
<div id="refs">
## Bibliography
</div>
